### 2013-09-12 paper_draft.tex
Added a few sentences, fixed some minor things

%\subsection*{Survey issues}
%
%As the survey was taken online by voluntary participants and was mainly spread in the personal genetics community, the results do not reflect the general population, but over-represent those people most likely to be interested in a project such as openSNP: customers of DTC genetic testing companies and people with a high interest in biology. 
%

@@

%Our survey has shown that 
As people are concerned about their privacy and fear that stakeholders like employers, insurance companies, governments or advertisers might misuse the information \cite{Wolinsky2005}, policy makers start to react to those changes by having introduced laws like the 
\textit{Genetic Information Non-Discrimination Act} (GINA) in the United States or the \emph{Gendiagnostikgesetz} (GenDG) in Germany to minimize the impact of
widely available genetic information. DTC genetic testing companies themselves also try to educate their customers about the risks of releasing genetic data \cite{Lee2009}. Neither GINA nor the GenDG offer a complete protection of genetic discrimination, as certain things, such as life insurances, are not covered by those laws.

@@

Another risk regarding data quality that should be kept in mind is a possible bias in data availability on openSNP: only a subset of people buy DTC genetic testing, from which an even smaller subset is willing to publish the results, which can potentially lead to skewed GWAS-results. 21 people, mainly from underrepresented demographics, have been offered free genotyping using funding provided by the Wikimedia Germany association in order to mitigate this bias.

>>>>

%\subsection*{Survey issues}
%
%As the survey was taken online by voluntary participants and was mainly spread in the personal genetics community, the results do not reflect the general population, but over-represent those people most likely to be interested in a project such as openSNP: customers of DTC genetic testing companies and people with a high interest in biology. 
%
Here, we present openSNP, a crowdsourced resource that enables customers of DTC testing companies to share their genotypings with researchers and receive new annotation for their genetic variants. We have shown that customers of DTC companies are willing to share their data at no cost to researchers around the world and are willing to annotate their data with phenotypes.

@@

%Our survey has shown that 
As people are concerned about their privacy and fear that stakeholders like employers, insurance companies, governments or advertisers might misuse the information \cite{Wolinsky2005}, policy makers start to react to those changes by having introduced laws like the 
\textit{Genetic Information Non-Discrimination Act} (GINA) in the United States or the \emph{Gendiagnostikgesetz} (GenDG) in Germany to minimize the impact of
widely available genetic information. DTC genetic testing companies themselves also try to create communities that help in educating their customers about the risks of releasing genetic data \cite{Lee2009}. Neither GINA nor the GenDG offer a complete protection of genetic discrimination, as certain things, such as life insurances, are not covered by those laws.

@@

Another risk regarding data quality that should be kept in mind is a possible bias in data availability on openSNP: only a subset of people buy DTC genetic testing, from which an even smaller subset is willing to publish the results, which can potentially lead to skewed GWAS-results. 21 people, mainly from underrepresented demographics, have been offered free genotypings using funding provided by the Wikimedia Germany association in order to mitigate this bias.
